---
title: "HurricanMaps"
author: "Bruce Mallory"
date: "11/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("maps","tmap","hurricaneexposuredata","tidyverse","drat", "magrittr", "sf", "dplyr")
```

### Read in all the data

```{r}
addRepo("geanders")
data("hurr_tracks")
data("rain")
county.longlat <- map_data("county")   
data(county.fips)

floyd_track <- hurr_tracks %>% filter(storm_id=="Floyd-1999")
floyd_rain <- rain %>% filter(storm_id=="Floyd-1999")

allison_track <- hurr_tracks %>% filter(storm_id=="Allison-2001")
allison_rain <- rain %>% filter(storm_id=="Allison-2001")

mp_states <- c("texas", "oklahoma", "kansas",
               "iowa", "missouri", "arkansas", "louisiana",
               "alabama", "mississippi", "georgia", "florida",
               "tennessee","kentucky", "indiana", 
               "wisconsin", "michigan", "illinois",
               "ohio", "west virginia", "pennsylvania",
               "south carolina", "north carolina", 
               "virginia", "delaware", "maryland",
               "new jersey", "district of columbia", 
               "new york", "connecticut", "rhode island",
               "massachusetts", "vermont", "new hampshire",
               "maine")

states <- map_data('state', region = mp_states)

```

### Calculate total rain/fip

```{r}
floyd_rain %<>% group_by(fips) %>% summarise(rain = sum(precip), .groups="drop")
allison_rain %<>% group_by(fips) %>% summarise(rain = sum(precip), .groups="drop")
```

### Join files to get a single file with long, lat, rain (which will be used to plot the data)

```{r}
#first split "polyname" into "region" and "subregion" so: county.fips and county.longlat can be joined
county.fips %<>%
  separate(polyname, c("region", "subregion"), "," )
county <- left_join(county.longlat,county.fips, by = c("region", "subregion"))
#second make "fips" numeric so: "county" and "xxx_rain" can be joined
floyd_rain$fips <- as.numeric(floyd_rain$fips)
floyd_rain <- left_join(floyd_rain, county, by = "fips")
allison_rain$fips <- as.numeric(allison_rain$fips)
allison_rain <- left_join(allison_rain, county, by = "fips")
```

### Now plot the data (first the rain, then the states, then the track) and format the graph

ggplot() +
  geom_polygon(data=floyd_rain, aes(x=long, y=lat, group=group, fill=rain), color="black", size=0.05) +
  geom_polygon(data=states, aes(x=long, y=lat, group=group), alpha=0, color="red")
  
ggplot() +
  geom_polygon(data=allison_rain, aes(x=long, y=lat, group=group, fill=rain), color="black", size=0.05)
  
  